---
title: "DATA 608 - Final Project: Beer data analysis"
author: "Peter Fernandes"

output: 
  html_document:
    #theme: cerulean
    #highlight: pygments
    #css: ./lab.css
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

For this project of DATA 608 I will be using the kaggle dataset <a href="https://www.kaggle.com/code/aieducation/how-delicious-is-this-beer-eda-lightgbm/data">beer_reviews.csv</a>. The dataset has 1.5 million beer reviews , with ratings for appearance, aroma, palate, taste, and overall impression.

# Objective

What are the factors that makes a beer favorite among the beer drinkers? How does the features like - beer taste, aroma, appearance, palate, review-time, beer style affects it's overall rating.

```{r warning=F, error=F, echo=F}
library(shiny)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(DataExplorer)
library(kableExtra)
library(tidymodels)
library(MASS)
library(sqldf)
library(ggplot2)
library(rsconnect)
```


# Dataset

## Columns description:

- `brewery_name`: The name of the brewery that made the beer. (String)
- `review_time`: The date and time of the review. (String)
- `review_overall`: The reviewer's overall rating of the beer on a scale of 1 to 5. (Float)
- `review_aroma`: The reviewer's rating of the beer's aroma on a scale of 1 to 5. (Float)
- `review_appearance`: The reviewer's rating of the beer's appearance on a scale of 1 to 5. (Float)
- `review_profilename`: The reviewer's username. (String)
- `beer_style`: The style of beer. (String)
- `review_palate`: The reviewer's rating of the beer's palate on a scale of 1 to 5. (Float)
- `review_taste`: The reviewer's rating of the beer's taste on a scale of 1 to 5. (Float)
- `beer_name`: The name of the beer. (String)
- `beer_abv`: The alcohol by volume of the beer. (Float)

## Loading of the dataset

I have downloaded the dataset from kaggale and placed in my local projects folder.

```{r}
df_beers <- read.csv('D:\\MSProjects\\608\\beer_reviews.csv',  stringsAsFactors=T)
```

The whole dataset consists of 1586614 observations and 14 variables.

```{r}
dim(df_beers)
```


```{r}
summary(df_beers)
```

## Cleaning of the data

We see from the summary of the dataset `beer_abv` has 67785 NA values. Since the dataset if large we will just drop all the NA records from our further analysis.

The new cleaned dataset has 1518829 observations.

```{r}
df_cleaned <- na.omit(df_beers)
dim(df_cleaned)
```

Next, we also exclude the column `index` its just a sequential number which is of no use for our analysis.


```{r}
df_cleaned <- df_cleaned[,-1]
```

Column `beer_style` has 2 values so this column will be split into 2 as `beer_style1` and `beer_style2` and `beer_style` column will be dropped from the dataset.

```{r}
df_cleaned <- df_cleaned %>% separate(beer_style, c("beer_style1","beer_style2"), extra="drop",fill = "right" )
```

```{r}
unique(as.factor(df_cleaned$beer_style1))
```


```{r}
head(df_cleaned,10) %>% kable() %>% kable_styling()
```
## Splitting of the dataset


I will split the dataset into 70% training and 30% testing dataset as `df_train` and `df_test` respectively.

```{r}
set.seed(3)
df_split <- initial_split(df_cleaned, prop = 0.7)
df_train <- training(df_split)
df_test <- testing(df_split)

```

Our training dataset has 1063180 observations.

```{r}
summary(df_train)
dim(df_train)
```
# Data visualization

In the below scatter plots of `brewery_id` versus `review_overall` and `beer_abv` versus `review_overall`, we do not see any relationship.

```{r}
plot(df_train$brewery_id, df_train$review_overall, main="Brewery/ Ratings",
   xlab="Brewery ", ylab="Rating ", pch=19)
```


```{r}
plot(df_train$beer_abv, df_train$review_overall, main="Beer abv/ Ratings",
   xlab="Beer abv ", ylab="Rating ", pch=19)
```

Next, I plotted the correlation matrix for all the features and we see a positive coorelation among the partial review factors.


```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning=FALSE}
DataExplorer::plot_correlation(data = df_train,type = "all",cor_args = list("use" = "pairwise.complete.obs"))
```

# Model building

I build a linear regression model by choosing the significant factors and we see the model performs well with high Fstat, smaller residual error and significant p-value.
Residual histogram is almost normally distributed, qq plot has most of the point falling under the line with some skewing at both edges

```{r echo=FALSE}
residualPlots <- function(model) {
  # Make sure a model was passed
  if (is.null(model)) {
    return
  }
  
  layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
  plot(residuals(model))
  hist(model[["residuals"]], freq = FALSE, breaks = "fd", main = "Residual Histogram",
       xlab = "Residuals",col="lightgreen")
  lines(density(model[["residuals"]], kernel = "ep"),col="blue", lwd=3)
  curve(dnorm(x,mean=mean(model[["residuals"]]), sd=sd(model[["residuals"]])), col="red", lwd=3, lty="dotted", add=T)
  qqnorm(model[["residuals"]], main = "Residual Q-Q plot")
  qqline(model[["residuals"]],col="red", lwd=3, lty="dotted")
  par(mfrow = c(1, 1))
}
```

```{r}
df_train <- dplyr::select_if(df_train,is.numeric)

model1 <- lm(review_overall ~ ., df_train)


(model1_summary <- summary(model1))

residualPlots(model1_summary)
```




# Conclusion

Review time, aroma, appearance, palate and taste form the major factors for the overall rating of a beer.



# Shiny app

**App1 - Beer aroma / Overall rating by brewery name** 


```{r}
Q1<-sqldf("
select 
review_overall, review_time, review_aroma ,brewery_name
from 
(
select 
            review_overall, review_time, review_aroma ,brewery_name
            from df_beers where brewery_name <>''
          
            group by brewery_name
) group by 4")
```




```{r}
ui <- fluidPage(
  sidebarPanel(
    selectInput(inputId = "Brewery", label = "Brewery name:",
                choices = levels(as.factor(unique(Q1$brewery_name)))
    ),
    helpText("Beer aroma / Overall rating by brewery name"),
    width = "auto"
  ),
  plotOutput("plot1")
)
server<- function(input, output) {
  
  output$plot1 <-renderPlot({
    
    SelectedCause <- input$Brewery
    
    ggplot(data=Q1[Q1$brewery_name == SelectedCause,]
           , aes(x = review_aroma, y = review_overall)) +
      labs(x="review_aroma", y="review_overall", 
           title = "review_aroma/review_overall",
           #subtitle = paste("Caused by", SelectedCause)
           )+   
      geom_bar(stat="identity", fill="steelblue") + 
      #geom_hline(aes(yintercept = mean(review_overall, na.rm = TRUE), linetype = "National Average"), col="red", lwd=1) +
      scale_linetype(name = NULL) +
      theme_bw()
    
  })
}
shinyApp(ui = ui, server = server)

```


**APP2 :  Review appearance / Overall rating by brewery name and Beer style**

```{r}
Q2<-sqldf("
select 
review_overall, review_time, review_aroma ,brewery_name, beer_style,review_appearance
from 
(
select 
            review_overall, review_time, review_aroma ,brewery_name, beer_style,review_appearance
            from df_beers where brewery_name <>''
          
            group by brewery_name
) group by 4")
```


```{r}
ui <- fluidPage(
  sidebarPanel(
    selectInput(inputId = "Brewery", label = "Brewery name:",
                choices = levels(as.factor(unique(Q2$brewery_name)))
    ),
    selectInput(inputId = "BeerStyle", label = "Beer Style:",
               choices = levels(as.factor(unique(Q2$beer_style)))
    ),
    helpText("Review appearance / Overall rating by brewery name and Beer style"),
    width = "auto"
  ),
  plotOutput("plot2")
)
server<- function(input, output) {
  
  output$plot2 <-renderPlot({
    
    SelectedCause <- input$Brewery
    SelectedState <- input$BeerStyle
    
    ggplot(data = Q2[Q2$brewery_name == SelectedCause & Q2$beer_style == SelectedState,]) +
      geom_bar(aes(x = review_appearance, weight = review_overall), fill = "steelblue") +
      labs(x="Review appearance", y = "Review overall", 
           title = "Review appearance / Overall rating by brewery name and Beer style"
          # ,subtitle = paste("Caused by", SelectedCause, "in", SelectedState)
           ) +
      #geom_line(aes(x = Year, y = N.Crude.Rate, linetype = "National Average"), col = "red", lwd = 1) +
      scale_linetype(name = NULL) +
      theme_bw()
    
     
    ggplot(data=Q1[Q1$brewery_name == SelectedCause,]
           , aes(x = review_aroma, y = review_overall)) +
      labs(x="review_aroma", y="review_overall", 
           title = "review_aroma/review_overall",
           #subtitle = paste("Caused by", SelectedCause)
           )+   
      geom_bar(stat="identity", fill="steelblue") + 
      #geom_hline(aes(yintercept = mean(review_overall, na.rm = TRUE), linetype = "National Average"), col="red", lwd=1) +
      scale_linetype(name = NULL) +
      theme_bw()
    
  })
}
shinyApp(ui = ui, server = server)
```


